---
title: "Woman's NBA Statistics for 2016"
author: "Angela McMahon and Mia Fitzgerald"
date: "3/27/2026"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)


```

## Introduction 
Our data set is a record of the Woman's NBA player stats for 2016. It is the basic stats of each woman's NBA player. This includes their team, position, number of games played and started in, minutes per game, and all shooting stats. They measured the number of free throws, field goals, 2 point, and 3 point attempts each player had along with all made shots and the percentages of these categories. This data set has a sample size of 174 different players. 
Link to data set: https://www.basketball-reference.com/wnba/years/2016_totals.html 

## Data Import 
To load our data set, the table was translated into a CSV and the data was copied and pasted. The data was then stored as WNBA_2016_stats. There was 2 variables which has duplicate columns, these were Games (G) and Minutes Played (MP). The duplicate columns were removed from the data. The columns for shooting percentages used the symbol which R does not accommodate. Those percentage symbols were changed into P's. It should be noted the percentages are recorded in decimal form in the data set. 

```{r}
library(readr)
WNBA_2016stats <- read_csv("WNBA_2016stats.csv")
WNBA_2016stats <- WNBA_2016stats%>% select(-G2, -MP2)

```


## Univariate Data 
### Number of Games Started 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=GS)) +
  geom_bar()+
  labs(x="Games Started")

WNBA_2016stats%>%
  summarize(
    Avg_Started = mean(GS),
    StDev_Started = sd(GS))%>% kable()

```

Here the category used was Games Started. This was measured as the number of games a player started in the given year (2016). On average a player starts 11.8 games in a year with a standard deviation of 13.9. It is interesting the highest count is when Games Started is zero, so many players do not start in a game. However, the second highest is over 30, indicating that players who start once will often continue to start. And those who do not will most likely not for most/all of the year. 

### Player Field Goal Percentage 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=FGP)) +
  geom_density()+
  labs(x="Field Goal Percentage")

WNBA_FGP_2016 <- WNBA_2016stats %>% 
  select(FGP)
summary(WNBA_FGP_2016) %>%
  kable()

```

The variable investigated here was player field goal percentage. These were recorded as field goal shots attempted and made then proportioned to find the percentage of shots made. Displayed in the density plot, the peak percentage on average for players is a little over 40%. The first quantile is around 36% and the third is about 48%, this represents 75% of the players average percentage in the data set. There is no extreme outliers present. 

### Number of Assists 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=AST))+
  geom_histogram(binwidth=5)+
  labs(x="Assists")

WNBA_2016assists <- WNBA_2016stats %>% 
  select(AST)
summary(WNBA_2016assists) %>% 
  kable()

```

The last variable examined in the univariate section is player assists in a given year. The histogram depicts the data is skewed to the right. There is one outlier right below 200 assists. The average number of assists for a WNBA player in 2016 is 42. While the median is 27, supporting the histogram showing a skewed relation. The first quantile is 9 and the third quantile is 66. 


## Bivariate Data 
### Teams and their Average Free Throw Percentage 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=Team, y=FTP))+
  geom_boxplot()+
  labs(x="WNBA Team", y="Free Throw Percentage")

WNBA_2016stats%>%
  group_by(Team) %>%
  summarize(
    Avg_FTP = mean(FTP, na.rm=TRUE),
    StDev_FTP = sd(FTP, na.rm=TRUE)) %>% 
  kable()

```

The two variables examined are WNBA teams and their free throw percentages. In the histogram, almost all teams have players with free throw percentages above 70%. With an exception of CON, while ATL and PHO boarder 70%. There are 6 teams with lower outliers. This includes ATL, CHI, CON, DAL, IND, and SAS. And one team with a high outlier which was WAS. It is impressive to see almost all WNBA teams are in the 80% range for player free throw percentage. With the standard deviations between 10-17% (0.10-0.17).


### Number of Free Throw Attmepts and Number Made 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=FTA, y=FT))+
  geom_point(alpha=0.3)+
  geom_smooth(method="lm")+
  labs(x="Free Throw Attempts", y="Free Throws Made")
           
```

Player free throw attempts and player free throw shots made have a positive, relatively strong correlation. Most of the points are concentrated between 0 and 25 on both axis. There are four outliers past 175. The strong correlation highly to these players shooting skills as most points are either touching or slightly off of the line of best fit. Although it does not show direct connection to the univariate views, it supports the above conclusions. 

## Trivariate Data 
### Player Position and the Number and Two and Three Point Shots Made 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=TwoPt, y= ThPt, color=Pos))+
  geom_point(alpha=0.5)+
  labs(x="Two Point Shots Made", y="Three Point Shots Made")

WNBA_2016stats%>%
  ggplot(aes(x=TwoPt, y=ThPt))+
  geom_point(alpha=0.3)+
  labs(x="Two Point Shots Made", y="Three Point Shots Made")+
  facet_wrap(~Pos)


```

The variables two-point shots made, three-point shots made, and position show a weak positive correlation in the first scatterplot. In the facetted scatterplots, we can see the same weak, positive correlation for the Forwards (F), Centers (C), and Guards (G) while the rotating players (F-G/F-C) show an even weaker positive to almost no correlation The first scatterplot we can see again the data is concentrated in the bottom left below 25. However, we can also notice the Guards and Forwards are making the most three point shots in a year. There are multiple outliers towards the top of the graph and a few on the right end as well. The faceted scatterplots support the conclusion guards and fowards are making the most 3 point shots. It also shows they are making on average more shots then the centers or rotating players.The center position shows an outlier on the far right, the same as noted earlier. 

## Conclusion 
The Woman's NBA stats for 2016 give lots of insight into the strengths and weaknesses of the teams and players. The data reveals a wide range of key roles of players including the shooting activity of guards and forwards. Over 60 players do not start but are still contibuting in other areas of the game including assists and different types of shots. Shooting percentages and assist numbers highlight both individual skills and teamwork. Both the bivariate and trivariate analyses demonstrate similar ideas. Factors like team, position, and shot selection interact to shape the outcomes on the court. Overall, the data set underscore the diversity of talent in the WBNA and the multifaceted nature of basketball games offering a comprehensive view of the 2016 basketball year.
