---
title: "Project 1"
author: "Angela McMahon and Mia Fitzgerald"
date: "3/27/2026"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)


```

## Introduction 
Our data set is a record of the Woman's NBA stats for 2016. It is the basic stats of each woman's NBA player. This includes their team, position, number of games played and started in, minutes per game, and all shooting stats. They measured the number of free throws, field goals, 2 point, and 3 point attempts each player had along with all made shots and the percentages of these categories. This data set has a sample size of 174 different players. 
Link to data set: https://www.basketball-reference.com/wnba/years/2016_totals.html 

## Data Import 
To load our data set, the table was translated into a CSV and the data was copied and pasted. The data was then stored as WNBA_2016_stats. There was 2 variables which has duplicate columns, these were Games (G) and Minutes Played (MP). The duplicate column was removed from the data. The columns for shooting percentages used the symbol which R does not accommodate. Those percentage symbols were changed into P's. 

```{r}
library(readr)
WNBA_2016stats <- read_csv("WNBA_2016stats.csv")
WNBA_2016stats <- WNBA_2016stats%>% select(-G2, -MP2)
#View(WNBA_2016stats)
```



## Univariate Data 

### Number of Games Started 
#games started 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=GS)) +
  geom_bar()+
  labs(x="Games Started")

WNBA_2016stats%>%
  summarize(
    Avg_Started = mean(GS),
    StDev_Started = sd(GS))%>% kable()

#another numeric summary? / fix??

```

Here the category used was Games Started. This was measured as the number of games a player started in the given year sample. On average a player starts 11.8 games in a year with a standard deviation of 13.9. It is interesting the highest count is when Games Started is zero, so many players do not start in a game. However, the second highest is over 30, indicating that players who start once will often continue to start. 

### Player Field Goal Percentage 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=FGP)) +
  geom_density()+
  labs(x="Field Goal Percentage")

#add numeric summaries 
```

#Several sentences discussing units, category meanings, interesting results, outliers, etc.



### Number of Assists 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=AST))+
  geom_histogram(binwidth=5)+
  labs(x="Assists")

# add numeric summaries 
```

#Several sentences discussing units, category meanings, interesting results, outliers, etc.



## Bivariate Data 

### Teams and their Average Free Throw Percentage 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=Team, y=FTP))+
  geom_boxplot()+
  labs(x="WNBA Team", y="Free Throw Percentage")

#numeric summarize broken down by category (team)
```

#A couple of sentences discussing any visible relationship and any additional insights that build on the univariate views of the data.



### Number of Free Throw Attmepts and Number Made 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=FTA, y=FT))+
  geom_point(alpha=0.3)+
  geom_smooth(method="lm")+
  labs(x="Free Throw Attempts", y="Free Throws Made")
           
```


#the correlation between the variables
#A couple of sentences discussing any visible relationship and any additional insights that build on the univariate views of the data.



## Trivariate Data 

### Player Position and the Number and Two and Three Point Shots Made 

```{r}
WNBA_2016stats%>%
  ggplot(aes(x=TwoPt, y= ThPt, color=Pos))+
  geom_point(alpha=0.5)+
  labs(x="Two Point Shots Made", y="Three Point Shots Made")

WNBA_2016stats%>%
  ggplot(aes(x=TwoPt, y=ThPt))+
  geom_point(alpha=0.3)+
  labs(x="Two Point Shots Made", y="Three Point Shots Made")+
  facet_wrap(~Pos)


```

#The correlation between the variables, broken out by the categories
#A couple of sentences discussing any visible relationship(s) and any additional insights that build on the earlier views of the data.


## Conclusion 


